---
title: "Info 201 Final Project-without code"
author: "Jessie Lu, Aaron Zheng"
date: "2023-12-08"
output: html_document
---

## Introduction

Los Angeles is a sprawling and diverse metropolis with a mix of neighborhoods representing various socio-economic and cultural backgrounds. In 2022, Los Angeles experienced an 11.6% increase in reported crimes compared to the previous year, marking the highest level of crime in the city in five years. At the same time, its rate of property crime was higher than the national rate. With California's overall crime rate at 83 incidents per square mile, Los Angeles notably exceeded this figure with a rate of 260 incidents per square mile. Moreover, crime reporting in Los Angeles saw a 4% increase from the previous year, indicating an increased number of reported incidents. 

### Why we interested in

The reason we chose this topic is also affected by my friend's stolen car this year when she went to LA to enjoy her summer vacation. At that time, she was in a shopping mall. Based on her memory, she did not put any valuable stuff in her car that may attract robbers. She called 911 as soon as possible and completed records of words. The police checked the monitor and told her to be patient. They will give her an answer. After four months, she received a reply from LA police officers that her car was stolen by a professional car theft gang. They usually appear in the area, and their goal is to have visitors with decent cars. Although she knew what happened, she told us that she still could not imagine how her car had been stolen within two hours.

### Research Questions

#### Our group wants to find out:

1. What the most common types of crimes
2. Which area has the highest rate of crime and arrest
3. Which gender and age are mostly involved in crimes.
4. Which gender and age are mostly involved in arrest.

### Datasets Choice

We collected two datasets for our project. The first one is crime reporting which allows for a deep exploration of how different communities are affected by and respond to criminal activities. Analyzing crime reporting data provides insights into the trends and patterns of criminal activities over time. The dataset of arrest reporting allows for a clear observation of the effectiveness of the LA police office. Researchers and law enforcement can identify which types of crimes are more prevalent in certain areas, helping to develop targeted intervention and prevention strategies. Assessing crime reports helps law enforcement agencies understand where resources are most needed and refine their approaches to crime prevention.
The second one is arrest reporting, which can assist law enforcement in contributing to the development of effective crime prevention strategies. Law enforcement can implement targeted initiatives to address root causes and deter criminal behavior. Additionally, Policymakers can use these datasets to inform decision-making related to public safety policies and resource allocation. Understanding crime hotspots and trends guides the allocation of resources for law enforcement, community programs, and social services. These datasets offer a comprehensive view of urban dynamics, community safety, law enforcement strategies, and societal responses to crime, which contribute to effective policies and the well-being of communities.

### Benefits of our audience

Our first dataset-crime reporting, reflects incidents of crime in the City of Los Angeles from 2010 - 2019, allowing for an in-depth analysis of trends, patterns, and changes in crime rates over a significant period. In this dataset, we have a large amount of effective information about victims' gender and age, reporting date, crime type, and crime area. Our second dataset, arrest reporting, reflects arrest incidents in the City of Los Angeles from 2010 to 2019, which includes the following columns: arrest date, crime type, crime area, gender, and age. These two datasets have similar columns, which is highly fit to combine. Analysis of them can help understand how crime evolves and the effectiveness of law enforcement strategies. Our target audience, Policy Makers and Government Officials, might seek insights to create or modify policies related to public safety, law enforcement funding, community programs, etc. These data can help visitors and residents improve.
For residents in LA, this dataset provides them with up-to-date and accurate information about the types and frequency of crimes in their neighborhoods. This awareness is crucial for understanding local safety conditions. Knowledge of criminal activities renders residents to take proactive measures for personal and community safety. This can include implementing neighborhood watch programs, securing homes, and being vigilant about suspicious activities. For visitors, they can use this data to plan their activities and choose accommodations in areas with lower crime rates. This helps ensure a safer and more enjoyable experience during their stay.


## Explaining dataset and Methods

### Load Packages used
```{r,echo=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(IRdisplay)
```

### Load original datasets-Crime reporting

```{r,echo=FALSE}
crime_2010_2019<-read_delim("C:/Users/11499/Desktop/info 201/project/Crime_Data_from_2010_to_2019.csv")
```

The crime reporting dataset includes 2,135,657 (2.13 millions) rows, and 28 columns, which reflects incidents of crime in the City of Los Angeles from 2010 - 2019, allowing for an in-depth analysis of trends, patterns, and changes in crime rates over a significant period. In this dataset, we have a large amount of effective information about reporting ID, victims' gender, descent and age, reporting date, crime type and area, and location. 
Moreover,our crime reporting dataset includes variables "DR Number","Date Reported","Time Occurred","Area Name", "Crime Code Description", "Victim Age", "Victim Sex", "Victim Descent", "Premise Description", "Status Description", "Location”. 

### Load original datasets-Arrest reporting

```{r,echo=FALSE}
arrest_2010_2019<-read_delim("C:/Users/11499/Desktop/info 201/project/Arrest_Data_from_2010_to_2019.csv")
```

Our arrest reporting dataset includes 1,320,983 rows and 25 columns, which reflects arrest incidents in the City of Los Angeles from 2010 to 2019, which also includes the following columns: arrest date, crime type, crime area, gender, age and so on. These two datasets have similar columns, which is highly fit to combine. Analysis of them can help understand how crime evolves and the effectiveness of law enforcement strategies. The arrest reporting dataset includes variables "Report ID", "Report Type", "Arrest Date", "Time", "Area Name", , "Age", "Sex Code", "Descent Code", "Charge Group Description", "Disposition Description", "Address", "Location". 
The dataset of crime reporting are transcribed from original crime reports that are typed on paper. It includes all the crimes that happened in Los Angeles. The dataset of arrest reporting is transcribed from original arrest reports that are typed on paper, and therefore, there may be some inaccuracies within the data. 
Besides, these two datasets are all collected by LAPD and published by LA City, which are non-federal datasets and are available for everyone without any extra fee. They are intended for public access and use and fall under the Creative Commons Public Domain Dedication (CC0), indicating minimal restrictions on access and use. We collected two of them from DATA.GOV in the form of a CSV file. Users can view datasets by the federal agency, state, city, or county that publishes them.

### Data select
#### Select the Columns that we need.

```{r,echo=FALSE}
crime_2010_2019 <- crime_2010_2019 %>% select("DR_NO", "Date Rptd", "TIME OCC", "AREA NAME", "Crm Cd Desc", "Vict Age", "Vict Sex", "Vict Descent", "Premis Desc", "Status Desc", "LOCATION")

arrest_2010_2019 <- arrest_2010_2019 %>% select("Report ID", "Report Type", "Arrest Date", "Time", "Area Name", , "Age", "Sex Code", "Descent Code", "Charge Group Description", "Disposition Description", "Address", "Location")
```

### Load Combined datasets

```{r,echo=FALSE}
# Combine data with the key report ID and DR_NO
arrest_2010_2019$"Report ID" <- as.character(arrest_2010_2019$"Report ID")

combine <- crime_2010_2019 %>% inner_join(arrest_2010_2019, by = c("DR_NO" = "Report ID"))
```

After combining, we added a new column named “Arrest day" as a numerical  variable in our combined dataset, which represents the difference between crime reporting and arrest reporting. Besides, we added a new categorical variable named “Sex Match”, which shows whether crime reporting sex and arresting report sex are same. From the results, we can conclude that the most popular type of crime happened on the same sex or different sex.

### Data Cleaning

```{r,echo=FALSE}
# Clean the data
combine <- na.omit(combine)
```

Clean all the missing value in the final combined dataset.

### Data quality issue

In our combined datasets, we have issues on report ID, which is our merge key to combine these two datasets. Each row in the crime reporting dataset and arrest reporting dataset are correlated with the same report ID. However, most of the rows are in the wrong report ID. Although this row from both two datasets are talking about same event, they are not matching. Finally, we filter out 1161 effective rows and 22 columns without any missing values in our final combined dataset.
Each row represents each crime that is being arrested by LAPD. These rows will contain the crime reporting date, arrest date, victims sex, age and area, and arrest sex, age and areas. In sum, we can have the relevant information about victims and criminals.
The relevant variables directly correlated to our following results are “Crime Type”, “Crime Area”, “Crime Age”, “Arrest Age”, “Crime Sex” and “Crime Sex”. There are many missing values in age that are coded in 0, and sex and descent that are coded in X. When we started to analyze our datasets to obtain results in different forms of tables and figures, we removed all of those missing values.


## Talk about results 

### Question 1: What is the most common types of crimes?

```{r,echo=FALSE}
## Find the top 5 of the crime type
crime_data <- combine %>% 
  group_by(`Crm Cd Desc`) %>% 
  summarise(Frequency = n()) %>% 
  arrange(desc(Frequency)) %>% 
  head(5)

crime_data %>%
  mutate(`Crm Cd Desc` = reorder(`Crm Cd Desc`, Frequency)) %>%
  ggplot(aes(x = "", y = Frequency, fill = `Crm Cd Desc`)) +
  geom_bar(stat = "identity", width = 5) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("red", "green", "purple", "yellow", "pink")) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Crime Type",title ="Top 5 Crime Type") +
  geom_text(aes(label = paste0(round((Frequency/sum(Frequency))*100, 1), "%")), position = position_stack(vjust = 0.5))


```

The pie shows the distribution of the top 5 types of crime from 2010 to 2019. Shoplifting - Petty ($950 & under) is the most common crime type, constituting 69% of the total crimes,which refers to the act of stealing goods from a retail store without paying for them. The color associated with this category on the chart is pink. Theft Plain - Petty ($950 & under), represented by yellow, constituting 28.6% of the total crimes. Battery - Simple Assault, represented in blue; this crime type constitutes 1.4% of total crimes. Intimate Partner - Simple Assault, shown in green, constitutes 0.94% of all crimes. Lastly, the other Miscellaneous Crime category is at 1.19%, shown in red on the chart.  

### Question 2: Which area has the highest rate of crime and arrest

```{r,echo=FALSE}
## Top 3 area that the crimes happened:
region_Top3<- head(sort(table(combine$`AREA NAME`), decreasing = TRUE), 3)

region_Top3 <- data.frame(region_Top3)

region_Top3 %>%
  mutate(Var1 = reorder(Var1, Freq)) %>%
  ggplot(aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("red", "yellow", "lightblue")) + 
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Region", title = "Top 3 Area that Crime Happened") +
  geom_text(aes(label = paste0(round((Freq/sum(Freq))*100, 1), "%")), position = position_stack(vjust = 0.5))
```

Topanga Region: The vast majority of crimes, 93.8%, occurred in the Topanga region, depicted in light blue. This region dominates the chart, indicating it is the area with the highest crime count in the dataset. Harbor Region, represented in yellow, this region accounts for a second large portion of the crimes, specifically 4.9%. West LA Region, this region has the third large visible share of crimes, shown in red, with a percentage of 1.3% portion of the crimes. The chart clearly indicates a significant concentration of crime in the Topanga region compared to the other two regions.

### Topanga total crime map

```{r,echo=FALSE}
##The Topanga total crime map
library(magick)
The_Topanga_Total_Crime_Map <- image_read("C:/Users/11499/Desktop/3.PNG.png")
plot(The_Topanga_Total_Crime_Map)
```

### Question 3: Which age are mostly involved in crimes

```{r,echo=FALSE}
## filter out all the victim ages equal to 0:
age_Data <- combine %>% filter(`Vict Age` != 0)

## make the plot
ggplot(age_Data, aes(x = `Vict Age`)) +
  geom_histogram(binwidth = 5, fill = "red", color = "black") +
  labs(x = "Victim Age", y = "Count", title ="Victim Age Distribution")
```

The victim's age ranges from 15 to 75, with a focus on 20 to 30. We assume that people in these pages are more likely to engage in risk-taking behavior or be involved in activities that can increase their exposure to crime.

### Question 4: Which age are mostly involved in arrest

```{r,echo=FALSE}
## filter out all the victim ages equal to 0:
age_Data <- combine %>% filter(`Age` != 0)

## make the plot
ggplot(age_Data, aes(x = `Age`)) +
  geom_histogram(binwidth = 5, fill = "red", color = "black") +
  labs(x = "Age", y = "Count", title ="Arrest Age Distribution")
```

For arresting age distribution, the age ranges from 15 to 75, which is similar to our previous victims age distribution. But it focuses mostly on 18 rather than 20, which is different from our previous distribution.

### Question 5: Which gender are most involved in crimes reporting  

```{r,echo=FALSE}
##filter out all the X value
sex_data <- combine %>%
  filter(`Vict Sex` != "X") %>%
  group_by(`Vict Sex`) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(percentage = count / sum(count) * 100)

## make the pie chart
  ggplot(sex_data, aes(x = "", y = percentage, fill = `Vict Sex`)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  labs(x = "", y = "Percentage", title = "Victim Sex Distribution", fill = "Victim Sex") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), position = position_stack(vjust = 0.5))
```

### Question 6: Which gender are most involved in arrest reporting

```{r,echo=FALSE}
## count for each sex and filter out all the X value
sex_data <- combine %>%
  filter(`Sex Code` != "X") %>%
  group_by(`Sex Code`) %>%
  summarise(count = n(), .groups = 'drop') %>%
  mutate(percentage = count / sum(count) * 100)

## make the pie chart
  ggplot(sex_data, aes(x = "", y = percentage, fill = `Sex Code`)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  labs(x = "", y = "Percentage", title = "Arrest Sex Distribution", fill = "Sex Code") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), position = position_stack(vjust = 0.5)) 
```

Our last result is about sex both in crime reporting and arrest reporting. We can notice that male is more likely to be victims than females, and have a lower rate of arrest than females as well. In these two pie charts we can see that the first one is Victims Sex Distribution, with 78.8% male and 21.2% female. The second one is Arrest Sex Distribution, with 38.7% male and 61.3% female.
Furthermore, we connect different columns in our combined dataset to produce additional graphs. 

### Victim sex in top crime area

```{r,echo=FALSE}
## Victim sex in top crime rate area
combine%>%
  filter(`AREA NAME` %in% c('Topanga','West LA','Harbor') & !`Vict Sex` %in% c('X','H'))%>%
  group_by(`AREA NAME`)%>%
  count(`Vict Sex`)%>%
  mutate(prop = n/sum(n))%>%
  ggplot(aes('',prop, fill=`Vict Sex`))+
  geom_bar(stat='identity', color = 'white')+
  labs(title='Victim Sex distribution',
       subtitle='Except Southeast, Male victims are more than female in other regions',
       x='Sex',
       y='Numbers of Victim Sex')+
  theme_bw()+
  coord_polar('y')+
  facet_wrap(~`AREA NAME`, nrow=1)
```

The first one is the image depicting three pie charts. Each represents the sex distribution of arrests in three different regions: Harbor, Topanga, and West LA. In Harbor, 64% of arrests were of males, while 36% were of females (depicted in red). In Topanga, the distribution is identical to Harbor, with males accounting for 62% of arrests and females 38%. West LA also shows similar distribution, with 71% of arrests being males and 29% females. From these pie charts, we are able to not only see clearly the distribution of sex in
these three region with highest crime rates, but also variation on sex.

## Arrest sex in top crime area

```{r,echo=FALSE}
## Arrest sex in top crime rate area
combine%>%
  filter(`AREA NAME` %in% c('Topanga','West LA','Harbor') & !`Vict Sex` %in% c('X','H'))%>%
  group_by(`AREA NAME`)%>%
  count(`Sex Code`)%>%
  mutate(prop = n/sum(n))%>%
  ggplot(aes('',prop, fill=`Sex Code`))+
  geom_bar(stat='identity', color = 'white')+
  labs(title='Arrest Sex distribution',
       x='Arrest Sex',
       y='Percentage in 3 top crime area')+
  theme_bw()+
  coord_polar('y')+
  facet_wrap(~`AREA NAME`, nrow=1)
```

The second graph is also three pie charts depicting three regions—Harbor, Topanga, and West LA, but illustrating the distribution of arrest by sex. The chart shows that in Harbor, the majority of victims are male (depicted in cyan), accounting for approximately 90%, while female victims (depicted in red) constitute about 10%. The distribution of Topanga with male victims at roughly 78% and female victims at about 22%. West LA, with female victims representing a larger proportion, around 28%, and male victims at approximately 72%.
From these pie charts, we are able to not only see clearly the distribution of arrest sex in these three region with highest crime rates, but also variation.

### Sex match

```{r,echo=FALSE}
# using the mutate function from dplyr
combine <- dplyr::mutate(combine, Sex_Match = ifelse(`Vict Sex` == `Sex Code`, "Same Sex", "Different Sex"))

# calculus the percentage
count_data <- combine %>%
                group_by(Sex_Match) %>%
                summarise(n = n()) %>%
                mutate(percentage = n / sum(n))

# plot the pie chart
ggplot(count_data, aes(x = "", y = n, fill = Sex_Match)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(percentage*100), "%")), 
            position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, 
       title = "Pie chart of Sex_Match") +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

The last one is the pie chart of "sex match", which corresponds to our new numerical column in our combined dataset. This pie chart refers to a comparison of gender related to crime reports and arrests. The chart is divided into two categories: Same Sex, shown in cyan, makes up the majority of the chart at 52%, which means the arrest sex is same as the reported one. This indicates that in over half of the cases, individuals involved were of the same sex. Different Sex, represented in red, this category comprises 48% of the chart. This shows that just under half of the cases involved individuals of different sexes.

## Findings

Our results reveals clearly that the most popular crime type is shoplifting, which belongs to a act of theft. The following highest crime type is also theft. Based on our dataset, we can analyze that in LA, theft appear in a high frequency. Also, according to the topanga crime map, we observe that the more crimes occur in the northeast parts of Topanga, CA: about 26 per year. The southwest part of Topanga has fewer cases of crime with only 4 in a typical year. In topanga, more issues arise with places like airports, parks, and schools. Major airports always look like high-crime locations due to the large number of people and the low population nearby. Parks and designated recreational areas have the same problem. Of Topanga’s 7,061 residents, few live near recreational areas. Because many people visit, crime rates may appear higher even for safe parks. Ultimately crime happens where people are, whether they live there or not. The law enforcement and police officer should attach more importance on this area and also the act of theft. Based on our results, residents are able to take proactive measures to keep their property safe.
Besides, compared to victims age distribution and arrest age distribution, we can find that victims' age distribute mainly on 20 to 30, while arrest age distributed mostly on 18. We assume that people in age 20 to 30 are more likely to engage in risk-taking behavior or be involved in activities that can increase their exposure to crime, and infer that people aged 18 has higher rate of crime, which is corresponding to the higher rate of arrest. The age of 18 represents a critical developmental stage where individuals often expereince transitions from adolescence to adulthood. These transitions, such as entering the workforce or pursuing higher education, can influence exposure to various social and environmental factors that may contribute to criminal behavior. 
For final sex distribution, it is a surprising result,since females are more likely to be victims generally. Thus,we suppose that because of our methods on selecting data, we only select people who commit a crime and are arrested by LA police. Thus, we can see that the arrest sex is higher on females than male, We may not have a relatively reliable result on gender here. 

### Limitation

Unfortunately, our combined dataset only contains about 1000 effecive rows, which is still not enough for our topic on crime. We are supposed to have a larger dataset to derive more reliable results. Moreover, what we achieved right now only offer basic information for law enforcement agencies, police officer and residents.

### Future Work with takeway message

All of outputs benefit our audience of the area where crime are most likely to occur, the victims sex that are most likely to be.For example, police officers could target topangda with more resource. Visitor could pay more attention on this area after viewing our dataset. 
Our target audience, Policy Makers and Government Officials, might seek insights to create or modify policies related to public safety, law enforcement funding, community programs, etc. This data can empower visitors and residents by providing them with information about the safety of their neighborhoods. It raises awareness and community engagement, enabling individuals to take proactive measures to enhance their safety and well-being.
